\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-23 europa-basics.tex
minor typos and jpg to PDF converstion

The next few sections will cover \eu in a bottom-up fashion, going
from the generic theoretical framework that underlies \eu, to the
specific representation, modeling and algorithms implemented by it.  A
brief introduction to Constraint Programming and Constraint-Based
Planning is provided followed by the specific representation that \eu
uses to undertake Constraint-Based Planning.  Subsequently, we provide
an overview of \eus architecture and how it can be embedded in a
planning application such as \rx. Finally, \eus modeling, inference
and search mechanisms are covered in detail so that the reader can
understand how \rx can perform deliberative planning and react to
execution events by embedding and customizing \eu.

\subsection{Automated Planning and \eu}
\label{sec:planningfound}

@@

\begin{figure} \centering
  \includegraphics[scale=0.35]{figs/shopping_europa.pdf}
  \caption{\small A potential plan for a robot shopping for groceries.}
\label{fig:shop:shopping}
\vskip+0.1cm
\end{figure}

To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of
planning we start with a simple example.

@@

At the core of the \rx framework, is the deliberation engine, \eu,
which has a rich legacy from NASA missions \cite{mus98,rajan00,
  jonsson00,aichang04, bresina05}. \eu is a versatile Constraint-based
Temporal Planner which continues to be deployed on a diverse set of
applications including recently, planning for the International Space
Station \cite{barreiro09}. We motivate this section with some problem
domains this planner can handle and then proceed to go into detail.

@@

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-NQueens1.jpg}
\caption{\small N-Queens solution generated by \eu}
\label{fig:nqueens-2}
\vskip+0.1cm
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an \texttt{RCPSP} \cite{Bruckera99}
  problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

\paragraph{} \textit{Planning}: The Shopping Agent Problem
\cite{russelnorvig} is a variation of the problem with which we started this chapter.
In this problem, an agent needs to purchase a set of products
(milk, drill, etc) that are available at specific locations
(supermarket, hardware store, etc) and the agent is subject to
temporal (must complete tasks by specific deadlines) and resource
(fuel, carrying capacity, etc) constraints. The agent needs to figure
out what actions need to be performed to find and acquire the required
items, as well as when to perform each of those
actions. Fig. \ref{fig:shopping-1} shows a solution produced by \eu
for such a problem instance.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-Shopping0.jpg}
\caption{\small A \eu solution to a shopping agent problem domain where
  the agent needs to buy Bananas, Milk and a Drill.}
\label{fig:shopping-1}
\vskip+0.1cm
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/constraint-net.pdf}
\caption{\small Variables and Constraints represented as a Constraint Network}
\label{fig:constraintnet}
\end{figure}

>>>>

The next few sections will cover \eu in a bottom-up fashion, going
from the generic theoretical framework that underlies \eue, to the
specific representation, modeling and algorithms implemented by it.  A
brief introduction to Constraint Programming and Constraint-Based
Planning is provided followed by the specific representation that \eu
uses to undertake Constraint-Based Planning.  Subsequently, we provide
an overview of \eus architecture and how it can be embedded in a
planning application such as \rxe. Finally, \eus modeling, inference
and search mechanisms are covered in detail so that the reader can
understand how \rx can perform deliberative planning and react to
execution events by embedding and customizing \eu.

\subsection{Automated Planning}
\label{sec:planningfound}

@@

\begin{figure} \centering
  \includegraphics[scale=0.35]{figs/shopping_europa.pdf}
  \caption{\small A potential plan for a robot shopping for groceries.}
\label{fig:shop:shopping}
\end{figure}

To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of
planning we start with an example.

@@

At the core of the \rx framework, is the deliberation engine, \eue,
which has a rich legacy from NASA missions \cite{mus98,rajan00,
  jonsson00,aichang04, bresina05}. \eu is a versatile Constraint-based
Temporal Planner which continues to be deployed on a diverse set of
applications including recently, planning for the International Space
Station \cite{barreiro09}. We motivate this section with some problem
domains this planner can handle and then proceed to go into detail.

@@

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-NQueens1.jpg}
\caption{\small N-Queens solution generated by \eu}
\label{fig:nqueens-2}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.25]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an \texttt{RCPSP} \cite{Bruckera99}
  problem.}
\label{fig:rcpsp-1}
\end{figure}

\paragraph{} \textit{Planning}: The Shopping Agent Problem
\cite{russelnorvig} is a variation of the problem with which we
started this chapter.  In this problem, an agent needs to purchase a
set of products (milk, drill, etc) that are available at specific
locations (supermarket, hardware store, etc). The agent is subject to
temporal (must complete tasks by specific deadlines) and resource
(fuel, carrying capacity, etc) constraints and it needs to synthesize
actions that need to be performed to find and acquire the required
items, as well as when to perform each of those actions.
Fig. \ref{fig:shopping-1} shows a solution produced by \eu for such a
problem instance.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/Example-Shopping0.jpg}
\caption{\small A \eu solution to a shopping agent problem domain where
  the agent needs to buy Bananas, Milk and a Drill.}
\label{fig:shopping-1}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[scale=0.3]{figs/constraint-net.pdf}
\caption{\small Variables and Constraints represented as a Constraint Network}
\label{fig:constraintnet}
\end{figure}

### 2012-02-23 intro.tex
minor typos and jpg to PDF converstion

This chapter is organized as follows. We start with some foundational
concepts in our Planning framework in Section~\ref{sec:concepts} which
includes a brief overview of AI Planning in
Section~\ref{sec:planningfound}. In Section~\ref{sec:basics} we go
deeper into the \eu framework itself, the basis of all deliberation
within \rx. A key portion is Section~\ref{sec:europa:pr} which details
the internal plan representation essential to planning within a
situated agent like \rx. Modeling (Section~\ref{sec:europa:modeling}),
Inference (Section~\ref{sec:europa:inference}) and Search
(Section~\ref{sec:europa:search}) further detail the important details
associated with \eu for the and relevant to how \rx works to the
reader. With these foundational elements out of the way, we transition
to \rx itself, in Section~\ref{sec:arch}, bring out architectural
details in Section~\ref{sec:arch:trex}, deal with the execution cycle
in Section~\ref{sec:arch:exec}, highlight how \rx deliberates in
Section~\ref{sec:arch:europa} finally delving deeper into the key
concept of synchronization in Section~\ref{sec:arch:synch}. Further
details on deliberation within \rx (Section~\ref{sec:arch:plan}) and
how planning and execution are interleaved are shown in
Section~\ref{sec:arch:intertwine}. We offer results from our
experiments at sea in Section~\ref{sec:results}, offer some insights
into the near-term future goals of our research in
Section~\ref{sec:future} and conclude with
Section~\ref{sec:conclusion}.

>>>>

This chapter is organized as follows. We start with some foundational
concepts in our Planning framework in Section~\ref{sec:concepts} which
includes a brief overview of AI Planning in
Section~\ref{sec:planningfound}. In Section~\ref{sec:basics} we go
deeper into the \eu framework itself, the basis of all deliberation
within \rxe. A key portion is Section~\ref{sec:europa:pr} which
details the internal plan representation essential to planning within
a situated agent like \rxe. Modeling
(Section~\ref{sec:europa:modeling}), Inference
(Section~\ref{sec:europa:inference}) and Search
(Section~\ref{sec:europa:search}) further detail the important details
associated with \eu for the and relevant to how \rx works to the
reader. With these foundational elements out of the way, we transition
to \rx itself, in Section~\ref{sec:arch}, bring out architectural
details in Section~\ref{sec:arch:trex}, deal with the execution cycle
in Section~\ref{sec:arch:exec}, highlight how \rx deliberates in
Section~\ref{sec:arch:europa} finally delving deeper into the key
concept of synchronization in Section~\ref{sec:arch:synch}. Further
details on deliberation within \rx (Section~\ref{sec:arch:plan}) and
how planning and execution are interleaved are shown in
Section~\ref{sec:arch:intertwine}. We offer results from our
experiments at sea in Section~\ref{sec:results}, offer some insights
into the near-term future goals of our research in
Section~\ref{sec:future} and conclude with
Section~\ref{sec:conclusion}.

### 2012-02-23 related.tex
minor typos and jpg to PDF converstion

Our system, \rx, the focus of this chapter, is an agent control
structure formally defined as a composition of coordinated control
loops where sensing, planning and acting are resulting from concurrent
deliberating tasks within a formal framework.  \rx was designed after
carefully evaluating lessons learned from \texttt{IDEA} to which it
owes substantially the notion of partitioned problem solving. However
partitioning in \rx is systematic and methodologically principled
where we manage the information flow within the partitioned structure
to ensure consistency in order to direct the flow of goals and
observations in a timely manner. The resulting control structure
improves scalability since many details of each controller can be
encapsulated within a single control loop.  Furthermore, partitioning
increases robustness since controller failure can be localized to
enable graceful system degradation, making this an effective
\emph{divide-and-conquer} approach to the overall control problem.

>>>>

Our system, \rxe, the focus of this chapter, is an agent control
structure formally defined as a composition of coordinated control
loops where sensing, planning and acting are resulting from concurrent
deliberating tasks within a formal framework.  \rx was designed after
carefully evaluating lessons learned from \texttt{IDEA} to which it
owes substantially the notion of partitioned problem solving. However
partitioning in \rx is systematic and methodologically principled
where we manage the information flow within the partitioned structure
to ensure consistency in order to direct the flow of goals and
observations in a timely manner. The resulting control structure
improves scalability since many details of each controller can be
encapsulated within a single control loop.  Furthermore, partitioning
increases robustness since controller failure can be localized to
enable graceful system degradation, making this an effective
\emph{divide-and-conquer} approach to the overall control problem.

### 2012-02-23 results.tex
minor typos and jpg to PDF converstion

The second innovation involved dynamic resource allocation associated
with the water samplers \cite{olaya11}. Using a priori sample
utilities coupled with temporal and spatial information maintained by
the deliberating planner in-situ, the algorithm dynamically generates
a policy on sample acquisition during the course of the mission. In so
doing, utility functions can be used to balance sparcity or abundance
of the feature encountered with the availability of samplers
dynamically. Coupled with periodic sub-sampled science data returned
by the vehicle when on the surface and goal-oriented commanding with
\rx, such a capability allows judicious use of the limited samplers by
a scientist from her desktop.

@@

Successive lagrangian experiments also followed a more innovative and
robotically choreographed ``ballet'' between two AUVs, the Dorado and
the LRAUV \cite{bellingham10} in October 2010 as part of another
series of \can field experiments. The slower vehicle with a
fluorometer tracked a targeted chlorophyll patch seeded by a
GPS-enabled drifter, sending periodic updates of the patch centroid to
shore via the \od following a setup similar to
Fig.\ref{fig:dta-setup}, to the Dorado running \rx.

@@

More interesting experiments are being carried out as a basis to track
more dynamic features such as ocean fronts \cite{fronts11} and
tracking more capable drifting platforms. In particular experiments
with a profiling CTD instrument, the \texttt{Wirewalker}
\cite{rainville01,pinkel11} are resulting in new modalities of making
Lagrangian and Eulerian observations. Fig. \ref{sep11-wirewalker}
shows one such experiment in September 2011 with a GPS-enabled
\texttt{Wirewalker} was tracked by the Dorado running \rx. We are
currently investigating spatio-temporal correlations \cite{gneiting07}
of CTD data obtained by the \texttt{Wirewalker} and tying it to the
observations made in the contextual environment by the Dorado's
sensors.

@@

These distinctions make some vehicles more or less suitable to address
certain science goals and suggest some degree of intelligent planning
in the assignment strategy. Further, even if partial plans can be
synthesized automatically, the need to capture qualitative 'intent'
and desires in the final outcome of an asset allocation strategy
drives us towards a human-in-the-loop approach for asset
allocation. The vast uncertainty and complex interaction between
unknown quantities in the ocean environment dictate that any
comprehensive experimental methodology must rely in some measure on
human reasoning and adaptability.  Furthermore, the multidisciplinary
nature of ocean experiments suggests collaboration between multiple
human actors. Human cognitive capability then to capture these
uncertainties, evaluate options, factor intent and synthesize data
needed for asset allocation becomes increasingly important. Therefore,
mixed-initiative approaches are viable as a means to apportion problem
solving and inference between human and computational approaches. A
viable example was shown in the command and control of the twin Mars
rovers in \cite{aichang04,bresina05,bresina05a}\footnote{The MAPGEN
  system uses the same \eu planner used in \rx. It continues to be
  used routinely to this day on the MER mission.}.

>>>>

The second innovation involved dynamic resource allocation associated
with the water samplers \cite{olaya11}. Using a priori sample
utilities coupled with temporal and spatial information maintained by
the deliberating planner in-situ, the algorithm dynamically generates
a policy on sample acquisition during the course of the mission. In so
doing, utility functions can be used to balance sparcity or abundance
of the feature encountered with the availability of samplers
dynamically. Coupled with periodic sub-sampled science data returned
by the vehicle when on the surface and goal-oriented commanding with
\rxe, such a capability allows judicious use of the limited samplers
by a scientist from her desktop.

@@

Successive lagrangian experiments also followed a more innovative and
robotically choreographed ``ballet'' between two AUVs, the Dorado and
the LRAUV \cite{bellingham10} in October 2010 as part of another
series of \can field experiments. The slower vehicle with a
fluorometer tracked a targeted chlorophyll patch seeded by a
GPS-enabled drifter, sending periodic updates of the patch centroid to
shore via the \od following a setup similar to
Fig.\ref{fig:dta-setup}, to the Dorado running \rxe.

@@

More interesting experiments are being carried out as a basis to track
more dynamic features such as ocean fronts \cite{fronts11} and
tracking more capable drifting platforms. In particular experiments
with a profiling CTD instrument, the \texttt{Wirewalker}
\cite{rainville01,pinkel11} are resulting in new modalities of making
Lagrangian and Eulerian observations. Fig. \ref{sep11-wirewalker}
shows one such experiment in September 2011 with a GPS-enabled
\texttt{Wirewalker} was tracked by the Dorado running \rxe. We are
currently investigating spatio-temporal correlations \cite{gneiting07}
of CTD data obtained by the \texttt{Wirewalker} and tying it to the
observations made in the contextual environment by the Dorado's
sensors.

@@

These distinctions make some vehicles more or less suitable to address
certain science goals and suggest some degree of intelligent planning
in the assignment strategy. Further, even if partial plans can be
synthesized automatically, the need to capture qualitative 'intent'
and desires in the final outcome of an asset allocation strategy
drives us towards a human-in-the-loop approach for asset
allocation. The vast uncertainty and complex interaction between
unknown quantities in the ocean environment dictate that any
comprehensive experimental methodology must rely in some measure on
human reasoning and adaptability.  Furthermore, the multidisciplinary
nature of ocean experiments suggests collaboration between multiple
human actors. Human cognitive capability then to capture these
uncertainties, evaluate options, factor intent and synthesize data
needed for asset allocation becomes increasingly important. Therefore,
mixed-initiative approaches are viable as a means to apportion problem
solving and inference between human and computational approaches. A
viable example was shown in the command and control of the twin Mars
rovers in \cite{aichang04,bresina05,bresina05a}\footnote{The MAPGEN
  system uses the same \eu planner used in \rxe. It continues to be
  used routinely to this day on the MER mission.}.
